---
title: "Aggregating Species Records"
author: "Danielle Quinn"
date: "16/10/2019"
output: html_document
---

**Goal**  
Aggregate species data to a lower taxonomic resolution.

**Data**  
Presence/absence data of chub and shiner species from five sites. Variable names are common names ending in `.chub` or `.shiner`.

```{r, warning = FALSE, message = FALSE}
# Load library
library(dplyr)
library(magrittr)

# Simulate data
mydata <- data.frame(sample = LETTERS[1:5],
                     lake.chub = c(1, 1, 0, 1, 0),
                     creek.chub = c(1, 0, 0, 0, 0),
                     unknown.chub = c(0, 1, 1, 1, 0),
                     golden.shiner = c(0, 0, 1, 0, 1),
                     common.shiner = c(0, 0, 0, 0, 1),
                     unknown.shiner = c(0, 0, 0, 0, 1))

mydata
```

**Build Function**  
Create a function `myfun` that applies `rowSums()` to columns that end with the string provided.  

```{r}
myfun <- function(match) {rowSums(mydata %>% select(ends_with(match)))}
```

**Apply Function**  
Loop through the two taxonomic groups (`chub` and `shiner`) that you want to aggregate into single columns. For each string, apply `myfun`, populate new named column with these values converted to presence (1) or absence (0). Remove the aggregated columns and update the `mydata` object.  

```{r}
for(i in c("chub", "shiner")) {
  mydata %<>%
    mutate(!!i := ifelse(myfun(i) > 0, 1, 0)) %>%
    select(-ends_with(paste0(".", i)))
}

mydata
```